High-level cognition during story listening is reflected in high-order dynamic correlations in neural activity patterns

### 2020-11-20 paper/main.tex
removed unused figure, corrected figure ordering in main text
edits and tweaks to main text, recompiled diff.tex

%\newcommand{\ttests}{S1}
\newcommand{\intact}{S1}
\newcommand{\para}{S2}
\newcommand{\word}{S3}
\newcommand{\rest}{S4}

@@

\section*{Introduction}
A central goal in cognitive neuroscience is to elucidate the
\textit{neural code}: the mapping between (a) mental states or
cognitive representations and (b) neural activity patterns. One means
of testing models of the neural code is to ask how accurately that
model is able to ``translate'' neural activity patterns into known (or
hypothesized) mental states or cognitive
representations~\citep[e.g.,][]{HaxbEtal01, NormEtal06, TongPrat12,
  MitcEtal08a, KamiTong05, NishEtal11, PereEtal18, HuthEtal12,
  HuthEtal16}.  Training decoding models on different types of neural
features (Fig.~\ref{fig:patterns}a) can also help to elucidate which
specific aspects of neural activity patterns are informative about
cognition-- and, by extension, which types of neural activity patterns
might compose the neural code.  For example, prior work has used
region of interest analyses to estimate the anatomical locations of
specific neural representations~\citep[e.g.,][]{EtzeEtal09}, or to
compare the relative contributions to the neural code of multivariate
activity patterns versus dynamic correlations between neural activity
patterns~\citep[e.g.,][]{MannEtal18, FongEtal19}.  An emerging theme
in this literature is that cognition is mediated by dynamic
interactions between brain structures~\citep{Gros88, Fris00, SporHone06, BassEtal06,
  Turk13, DemeEtal19, SoloEtal19, LuriEtal18, PretEtal17, ZouEtal19,
  MackEtal17}.

\begin{figure}[tp]
  \centering
  \includegraphics[width=\textwidth]{figs/patterns}
  \caption{\textbf{Neural patterns.  a. A space of neural features.}
    Within-brain analyses are carried out within a single brain,
    whereas across-brain analyses compare neural patterns across two
    or more individuals' brains.  Univariate analyses characterize the
    activities of individual units (e.g., nodes, small networks,
    hierarchies of networks, etc.), whereas multivariate analyses
    characterize the patterns of activities across units.  Order 0
    patterns involve individual nodes; order 1 patterns involve
    node-node interactions; order 2 (and higher) patterns relate to
    interactions between homologous networks.  Each of these patterns
    may be static (e.g., averaging over time) or dynamic.
    \textbf{b. Summarizing neural patterns.}  To efficiently compute
    with complex neural patterns, it can be useful to characterize the
    patterns using summary measures.  Dimensionality reduction
    algorithms project the patterns onto lower-dimensional spaces
    whose dimensions reflect weighted combinations or non-linear
    transformations of the dimensions in the original space.  Graph
    measures characterize each unit's participation in its associated
    network.}
  \label{fig:patterns}
\end{figure}

@@

To gain insights into the above aspects of the neural code, we
developed a computational framework for estimating dynamic high-order
correlations in timeseries data. This framework provides an important
advance, in that it enables us to examine patterns of higher-order
correlations that are computationally intractable to estimate via
conventional methods.  Given a multivariate timeseries, our framework
provides timepoint-by-timepoint estimates of the first-order
correlations, second-order correlations, and so on.  Our approach
combines a kernel-based method for computing dynamic correlations in
timeseries data with a dimensionality reduction step
(Fig.~\ref{fig:patterns}b) that projects the resulting dynamic
correlations into a low-dimensional space.  We explored two
dimensionality reduction approaches: principle components
analysis~\citep[PCA;][]{Pear01}, which preserves an approximately
invertible transformation back to the original data~\citep[e.g., this
follows related approaches taken by][]{McInJirs19, TokeSomm19,
  GonzEtal19}; and a second non-invertible algorithm that explored
patterns in eigenvector centrality~\citep{Land95}.  This latter
approach characterizes correlations between each feature dimension's
relative \textit{position} in the network in favor of the specific
activity histories of different features~\citep[also
see][]{BetzEtal19, SizeEtal18, ReimEtal17}.

We validated our approach using synthetic data where the underlying
correlations were known.  We then applied our framework to a
neuroimaging dataset collected as participants listened to either an
audio recording of a ten-minute story, listened to a temporally
scrambled version of the story, or underwent a resting state
scan~\citep{SimoEtal16}.  Temporal scrambling has been used in a
growing number of studies, largely by Uri Hasson's group, to identify
brain regions that are sensitive to higher-order and longer-timescale
information (e.g., cross-sensory integration, rich narrative meaning,
complex situations, etc.) versus regions that are primarily sensitive
to low-order (e.g., sensory) information.  For example,
\cite{HassEtal08} argues that when brain areas are sensitive to fine
versus coarse temporal scrambling, this indicates that they are
``higher order'' in the sense that they process contextual information
pertaining to further-away timescales.  By contrast, low-level
regions, such as primary sensory cortices, do not meaningfully change
their responses (after correcting for presentation order) even when
the stimulus is scrambled at fine timescales.

@@

\subsection*{Cognitively relevant dynamic high-order correlations in
  fMRI data}
We used across-participant temporal decoders to identify cognitively
relevant neural patterns in fMRI data (see \textit{Forward inference
  and decoding accuracy}).  The dataset we examined~\citep[collected
by][]{SimoEtal16} comprised four experimental conditions that exposed
participants to stimuli that varied systematically in how cognitively
engaging they were.  The \textit{intact} experimental condition had
participants listen to an audio recording of a 10-minute story.  The
\textit{paragraph}-scrambled experimental condition had participants
listen to a temporally scrambled version of the story, where the
paragraphs occurred out of order (but where the same total set of
paragraphs were presented over the full listening interval).  All
participants in this condition experienced the scrambled paragraphs in
the same order.  The \textit{word}-scrambled experimental condition
had participants listen to a temporally scrambled version of the story
where the words in the story occurred in a random order.  All
participants in the word condition experienced the scrambled words in
the same order.  Finally, in a \textit{rest} experimental condition,
participants lay in the scanner with no overt stimulus, with their
eyes open (blinking as needed).  This dataset provided a convenient
means of testing our hypothesis that different levels of cognitive
processing and engagement are reflected in different orders of brain
activity dynamics.

\begin{figure}[tp]
  \centering
  \includegraphics[width=\textwidth]{figs/decode_level}
  \caption{\textbf{Across-participant timepoint decoding accuracy varies with
      correlation order and cognitive engagement.}
    \textbf{a.~Decoding accuracy as a function of order: PCA.}
    \textit{Order} ($x$-axis) refers to the maximum order of dynamic
    correlations that were available to the classifiers (see
    \textit{Feature weighting and testing}).  The reported
    across-participant decoding accuracies are averaged over all
    kernel shapes and widths (see \textit{Identifying robust decoding
      results}).  The $y$-values are displayed relative to chance
    accuracy (intact: $\frac{1}{300}$; paragraph: $\frac{1}{272}$;
    word: $\frac{1}{300}$; rest: $\frac{1}{400}$).  The error ribbons
    denote 95\% confidence intervals across cross-validation folds
    (i.e., random assignments of participants to the training and test
    sets).  The colors denote the experimental condition.  Arrows
    denote sets of features that yielded reliably higher (upwards
    facing) or lower (downward facing) decoding accuracy than the mean
    of all other features (via a two-tailed test, thresholded at
    $p < 0.05$).  Figure~\ref{fig:ttests} displays additional
    comparisons between the decoding accuracies achieved using
    different sets of neural features.  The circled values represent
    the maximum decoding accuracy within each experimental condition.
    \textbf{b.~Normalized timepoint decoding accuracy as a function of order:
      PCA.}  This panel displays the same results as Panel a, but here
    each curve has been normalized to have a maximum value of 1 and a
    minimum value of 0 (including the upper and lower bounds of the
    respective 95\% confidence intervals).  Panels a and b used PCA to
    project each high-dimensional pattern of dynamic correlations onto
    a lower-dimensional space.  \textbf{c.~Timepoint decoding accuracy as a
      function of order: eigenvector centrality.} This panel is in the
    same format as Panel a, but here eigenvector centrality has been
    used to project the high-dimensional patterns of dynamic
    correlations onto a lower-dimensional space.
    \textbf{d.~Normalized timepoint decoding accuracy as a function of order:
      eigenvector centrality.} This panel is in the same format as
    Panel b, but here eigenvector centrality has been used to project
    the high-dimensional patterns of dynamic correlations onto a
    lower-dimensional space.}
  \label{fig:decoding}
\end{figure}

@@

\begin{figure}[tp]
\centering
\includegraphics[width=1\textwidth]{figs/stats_heatmaps}
\caption{\textbf{Statistical summary of decoding accuracies for
    different neural features.}  Each column of matrices displays decoding
  results for one experimental condition (intact, paragraph, word, and
  rest).  We considered dynamic activity patterns (order 0) and
  dynamic correlations at different orders (order $> 0$).  We used
  two-tailed $t$-tests to compare the
  distributions of decoding accuracies obtained using each pair of
  features.  The distributions for each feature reflect the set of
  average decoding accuracies (across all kernel parameters), obtained
  for each random assignment of training and test groups.
  In the upper triangles of each matrix, warmer colors (positive $t$-values) indicate that the
  neural feature indicated in the given row yielded higher accuracy than the
  feature indicated in the given column.  Cooler colors (negative
  $t$-values) indicate that
  the feature in the given row yielded lower decoding accuracy than
  the feature in the given column.  The lower triangles of each map
  denote the corresponding $p$-values for the $t$-tests.  The diagonal
  entries display the relative average optimized weight given to each type of feature, in
  a decoder that included all feature types (see \textit{Feature
    weighting and testing}).}
\label{fig:ttests}
\end{figure}

Both sets of temporal decoding analyses yielded qualitatively similar
results for the auditory (non-rest) conditions of the experiment
(Fig.~\ref{fig:decoding}: pink, green, and teal lines;
Fig.~\ref{fig:ttests}: three leftmost columns).  The highest
decoding accuracy for participants who listened to the intact
(unscrambled) story was achieved using high-order dynamic correlations
(PCA: second-order; eigenvector-centrality: fourth-order).  Scrambled
versions of the story were best decoded by lower-order correlations
(PCA/paragraph: first-order; PCA/word: order zero; eigenvector
centrality/paragraph: order zero; eigenvector centrality/word: order
zero).  The two sets of analyses yielded different decoding results on
resting state data (Fig.~\ref{fig:decoding}: purple lines;
Fig.~\ref{fig:ttests}: rightmost column).  We note
that while the resting state times could be decoded reliably, the
accuracies were only very slightly above chance.  We speculate that
the decoders might have picked up on attentional drift, boredom, or
tiredness; we hypothesize that these all increased throughout the resting
state scan.  The decoders might be picking up on aspects of these
loosely defined cognitive states that are common across individuals.
The PCA-based approach achieved the highest resting state decoding
accuracy using order zero features (non-correlational,
activation-based), whereas the eigenvector centrality-based approach
achieved the highest resting state decoding accuracy using
second-order correlations.  Taken together, these analyses indicate
that high-level cognitive processing (while listening to the intact
story) is reflected in the dynamics of high-order correlations in
brain activity, whereas lower-level cognitive processing (while
listening to scrambled versions of the story that lack rich meaning)
is reflected in the dynamics of lower-order correlations and
non-correlational activity dynamics.  Further, these patterns are
associated both with the underlying activity patterns (characterized
using PCA) and also with the changing relative positions that
different brain areas occupy in their associated networks
(characterized using eigenvector centrality).

\begin{figure}[tp]
  \centering
  \includegraphics[width=\textwidth]{figs/most_abs}
  \caption{\textbf{Top terms associated with the endpoints of the
      strongest correlations.}  Each color corresponds to one order of
    inter-subject functional correlations.  To calculate the dynamic
    correlations, eigenvector centrality has been used to project
    the high-dimensional patterns of dynamic correlations onto a
    lower-dimensional space at each previous order, which allows us
    the map the brain regions at each order by retaining the
    features of the original space. The inflated brain plots
    display the locations of the endpoints of the 10 strongest
    (absolute value) correlations at each order, thresholded at 0.999,
    and projected onto the cortical surface~\citep{CombEtal19}.  The
    lists of terms on the right display the top five Neurosynth
    terms~\citep{RubiEtal17} decoded from the corresponding brain maps
    for each order.  Each row displays data from a different
    experimental condition.  Additional maps and their corresponding
    Neurosynth terms may be found in the \textit{Supplementary
      materials} (intact: Fig.~\intact; paragraph: Fig.~\para; word:
    Fig.~\word; rest: Fig.~\rest).}
  \label{fig:neurosynth}
\end{figure}

Having established that patterns of high-order correlations are
informative to decoders, we next wondered which specific networks of
brain regions contributed most to these patterns.  As a representative
example, we selected the kernel parameters that yielded decoding
accuracies that best matched the average accuracies across all of the
kernel parameters we examined.  Using Figure~\ref{fig:decoding}c as a
template, the best-matching kernel was a Laplace kernel with a width
of 50 (Fig.~\ref{fig:kernels}d).  We used this kernel to compute a
single $K$ by $K$ $n^\mathrm{th}$-order DISFC matrix for each
experimental condition.  We then used Neurosynth~\citep{RubiEtal17} to
compute the terms most highly associated with the most strongly
correlated pairs of regions in each of these matrices
(Fig.~\ref{fig:neurosynth}; see \textit{Reverse inference}).

For all of the story listening conditions (intact, paragraph, and
word), we found that first- and second-order correlations were most
strongly associated with auditory and speech processing areas.  During
intact story listening, third-order correlations reflected integration
with visual areas, and fourth-order correlations reflected integration
with areas associated with high-level cognition and cognitive control,
such as the ventrolateral prefrontal cortex.  However, during
listening to temporally scrambled stories, these higher-order
correlations instead involved interactions with additional regions
associated with speech and semantic processing.  By contrast, we found
a much different set of patterns in the resting state data.
First-order resting state correlations were most strongly associated
with regions involved in counting and numerical understanding.
Second-order resting state correlations were strongest in visual
areas; third-order correlations were strongest in task-positive areas;
and fourth-order correlations were strongest in regions associated
with autobiographical and episodic memory.  We carried out analogous
analyses to create maps (and decode the top associated Neurosynth
terms) for up to fifteenth-order correlations (Figs.~\intact, \para, \word, and
\rest).  Of note, examining fifteenth-order correlations between 700
nodes using conventional methods would have required storing roughly
$\frac{700^{2 \times 15}}{2} \approx 1.13 \times 10^{85}$ floating
point numbers-- assuming single-precision (32 bits each), this would
require roughly 32 times as many bits as there are molecules in the
known universe!  Although these fifteenth-order correlations do appear
(visually) to have some well-formed structure, we provide this latter
example primarily as a demonstration of the efficiency and scalability
of our approach.

@@

The notion that cognition is reflected in (and possibly mediated by)
patterns of first-order network dynamics has been suggested by or
proposed in myriad empirical studies and
reviews~\citep[e.g.,][]{DemeEtal19, Turk13, LuriEtal18, FongEtal19,
  ParkEtal18b, PretEtal17, MannEtal18, RoyEtal19, LiegEtal19, ZouEtal19,
  ChanGlov10, GonzEtal19}.  Our study extends this line of work by
finding cognitively relevant \textit{higher-order} network dynamics
that reflect ongoing cognition.  Our findings complement other work
that uses graph theory and topology to characterize how brain networks
reconfigure during cognition~\citep[e.g.,][]{BassEtal06, ZhenEtal19,
  McInJirs19, TokeSomm19, SizeEtal18, ReimEtal17, BetzEtal19}.

An open question not addressed by our study pertains to how different
structures integrate incoming information with different time
constants.  For example, one line of work suggests that the cortical
surface comprises a structured map such that nearby brain structures
process incoming information at similar timescales.  Low-level sensory
areas integrate information relatively quickly, whereas higher-level
regions integrate information relatively slowly~\citep{BaldEtal17,
  HassEtal08, HassEtal15, HoneEtal12a, LernEtal11, LernEtal14,
  ChieHone19}.  A similar hierarchy appears to play a role in
predicting future events~\citep{LeeEtal20}.  Other
related work in human and mouse brains indicates that the temporal
response profile of a given brain structure may relate to how strongly
connected that structure is with other brain
areas~\citep{FallEtal19}.  Further study is needed to understand the
role of temporal integration at different scales of network
interaction, and across different anatomical structures. Importantly,
our analyses do not speak to the physiological basis of higher-order
dynamics, and could reflect nonlinearities, chaotic patterns,
non-stationarities, or multistability, etc. However, our decoding analyses do indicate
that higher-order dynamics are
consistent across individuals, and therefore unlikely to be driven by
non-stimulus-driven dynamics which are unlikely to be similar across
individuals.

@@

Identifying high-order network dynamics associated with high-level
cognition required several important methods advances.  First, we used
kernel-based dynamic correlations to extended the notion of (static)
inter-subject functional connectivity~\citep{SimoEtal16} to a dynamic
measure of inter-subject functional connectivity (DISFC) that does not
rely on sliding windows~\citep[e.g., as in][]{MannEtal18}, and that
may be computed at individual timepoints.  This allowed us to
precisely characterize stimulus-evoked network dynamics that were
similar across individuals.  Second, we developed a computational
framework for efficiently and scalably estimating high-order dynamic
correlations.  Our approach uses dimensionality reduction algorithms
and graph measures to obtain low-dimensional embeddings of patterns of
network dynamics.  Third, we developed an analysis framework for
identifying robust decoding results by carrying out our analyses using
a range of parameter values and then identifying which results were
robust to specific parameter choices.

\subsection*{Concluding remarks}
Methods for characterizing the dynamics of high-order correlations in
neural data provide a window into the neural basis of cognition.  By
showing that high-level cognition is reflected in high-order network
dynamics, we have elucidated the next step on the path towards
understanding the neural basis of cognition.

>>>>

\newcommand{\kernelshape}{S1}
\newcommand{\kernelwidth}{S2}
\newcommand{\intact}{S3}
\newcommand{\para}{S4}
\newcommand{\word}{S5}
\newcommand{\rest}{S6}
\newcommand{\pca}{S7}

@@

\section*{Introduction}
A central goal in cognitive neuroscience is to elucidate the
\textit{neural code}: the mapping between (a) mental states or
cognitive representations and (b) neural activity patterns. One means
of testing models of the neural code is to ask how accurately that
model is able to ``translate'' neural activity patterns into known (or
hypothesized) mental states or cognitive
representations~\citep[e.g.,][]{HaxbEtal01, NormEtal06, TongPrat12,
  MitcEtal08a, KamiTong05, NishEtal11, PereEtal18, HuthEtal12,
  HuthEtal16}.  Training decoding models on different types of neural
features (Fig.~\ref{fig:patterns}a) can also help to elucidate which
specific aspects of neural activity patterns are informative about
cognition and, by extension, which types of neural activity patterns
might compose the neural code.  For example, prior work has used
region of interest analyses to estimate the anatomical locations of
specific neural representations~\citep[e.g.,][]{EtzeEtal09}, or to
compare the relative contributions to the neural code of multivariate
activity patterns versus dynamic correlations between neural activity
patterns~\citep[e.g.,][]{MannEtal18, FongEtal19}.  An emerging theme
in this literature is that cognition is mediated by dynamic
interactions between brain structures~\citep{Gros88, Fris00, SporHone06, BassEtal06,
  Turk13, DemeEtal19, SoloEtal19, LuriEtal18, PretEtal17, ZouEtal19,
  MackEtal17}.

\begin{figure}[tp]
  \centering
  \includegraphics[width=\textwidth]{figs/patterns}
  \caption{\textbf{Neural patterns.  a. A space of neural features.}
    Within-brain analyses are carried out within a single brain,
    whereas across-brain analyses compare neural patterns across two
    or more individuals' brains.  Univariate analyses characterize the
    activities of individual units (e.g., nodes, small networks,
    hierarchies of networks, etc.), whereas multivariate analyses
    characterize the patterns of activity across units.  Order 0
    patterns involve individual nodes; order 1 patterns involve
    node-node interactions; order 2 (and higher) patterns relate to
    interactions between homologous networks.  Each of these patterns
    may be static (e.g., averaging over time) or dynamic.
    \textbf{b. Summarizing neural patterns.}  To efficiently compute
    with complex neural patterns, it can be useful to characterize the
    patterns using summary measures.  Dimensionality reduction
    algorithms project the patterns onto lower-dimensional spaces
    whose dimensions reflect weighted combinations or non-linear
    transformations of the dimensions in the original space.  Graph
    measures characterize each unit's participation in its associated
    network.}
  \label{fig:patterns}
\end{figure}

@@

To gain insights into the above aspects of the neural code, we
developed a computational framework for estimating dynamic high-order
correlations in timeseries data. This framework provides an important
advance, in that it enables us to examine patterns of higher-order
correlations that are computationally intractable to estimate via
conventional methods.  Given a multivariate timeseries, our framework
provides timepoint-by-timepoint estimates of the first-order
correlations, second-order correlations, and so on.  Our approach
combines a kernel-based method for computing dynamic correlations in
timeseries data with a dimensionality reduction step
(Fig.~\ref{fig:patterns}b) that projects the resulting dynamic
correlations into a low-dimensional space.  We explored two
dimensionality reduction approaches: principle components
analysis~\citep[PCA;][]{Pear01}, which preserves an approximately
invertible transformation back to the original data~\citep[e.g., this
follows related approaches taken by][]{McInJirs19, TokeSomm19,
  GonzEtal19}; and a second non-invertible algorithm for computing
dynamic patterns in eigenvector centrality~\citep{Land95}.  This
latter approach characterizes correlations between each feature
dimension's relative \textit{position} in the network (at each moment
in time) in favor of the specific activity histories of different
features~\citep[also see][]{BetzEtal19, SizeEtal18, ReimEtal17}.

We validated our approach using synthetic data where the underlying
correlations were known.  We then applied our framework to a
neuroimaging dataset collected as participants listened to either an
audio recording of a ten-minute story, listened to a temporally
scrambled version of the story, or underwent a resting state
scan~\citep{SimoEtal16}.  Temporal scrambling has been used in a
growing number of studies, largely by Uri Hasson's group, to identify
brain regions that are sensitive to higher-order and longer-timescale
information (e.g., cross-sensory integration, rich narrative meaning,
complex situations, etc.) versus regions that are primarily sensitive
to low-order (e.g., sensory) information.  For example,
\cite{HassEtal08} argues that when brain areas are sensitive to fine
versus coarse temporal scrambling, this indicates that they are
``higher order'' in the sense that they process contextual information
pertaining to further-away timepoints.  By contrast, low-level
regions, such as primary sensory cortices, do not meaningfully change
their responses (after correcting for presentation order) even when
the stimulus is scrambled at fine timescales.

@@

\subsection*{Cognitively relevant dynamic high-order correlations in
  fMRI data}
We used across-participant temporal decoders to identify cognitively
relevant neural patterns in fMRI data (see \textit{Forward inference
  and decoding accuracy}).  The dataset we examined~\citep[collected
by][]{SimoEtal16} comprised four experimental conditions that exposed
participants to stimuli that varied systematically in how cognitively
engaging they were.  The \textit{intact} experimental condition had
participants listen to an audio recording of a 10-minute story.  The
\textit{paragraph}-scrambled experimental condition had participants
listen to a temporally scrambled version of the story, where the
paragraphs occurred out of order (but where the same total set of
paragraphs were presented over the full listening interval).  All
participants in this condition experienced the scrambled paragraphs in
the same order.  The \textit{word}-scrambled experimental condition
had participants listen to a temporally scrambled version of the story
where the words in the story occurred in a random order.  All
participants in the word condition experienced the scrambled words in
the same order.  Finally, in a \textit{rest} experimental condition,
participants lay in the scanner with no overt stimulus, with their
eyes open (blinking as needed).  This public dataset provided a convenient
means of testing our hypothesis that different levels of cognitive
processing and engagement are reflected in different orders of brain
activity dynamics.

\begin{figure}[tp]
  \centering
  \includegraphics[width=\textwidth]{figs/decode_level}
  \caption{\textbf{Across-participant timepoint decoding accuracy varies with
      correlation order and cognitive engagement.}
    \textbf{a.~Decoding accuracy as a function of order: PCA.}
    \textit{Order} ($x$-axis) refers to the maximum order of dynamic
    correlations that were available to the classifiers (see
    \textit{Feature weighting and testing}).  The reported
    across-participant decoding accuracies are averaged over all
    kernel shapes and widths (see \textit{Identifying robust decoding
      results}).  The $y$-values are displayed relative to chance
    accuracy (intact: $\frac{1}{300}$; paragraph: $\frac{1}{272}$;
    word: $\frac{1}{300}$; rest: $\frac{1}{400}$; these chance accuracies
    were subtracted from the observed accuracies to obtain the
    relative accuracies reported on the $y$-axis).  The error ribbons
    denote 95\% confidence intervals across cross-validation folds
    (i.e., random assignments of participants to the training and test
    sets).  The colors denote the experimental condition.  Arrows
    denote sets of features that yielded reliably higher (upwards
    facing) or lower (downward facing) decoding accuracy than the mean
    of all other features (via a two-tailed $t$-test, thresholded at
    $p < 0.05$).  Figure~\ref{fig:ttests} displays additional
    comparisons between the decoding accuracies achieved using
    different sets of neural features.  The circled values represent
    the maximum decoding accuracy within each experimental condition.
    \textbf{b.~Normalized timepoint decoding accuracy as a function of order:
      PCA.}  This panel displays the same results as Panel a, but here
    each curve has been normalized to have a maximum value of 1 and a
    minimum value of 0 (including the upper and lower bounds of the
    respective 95\% confidence intervals).  Panels a and b used PCA to
    project each high-dimensional pattern of dynamic correlations onto
    a lower-dimensional space.  \textbf{c.~Timepoint decoding accuracy as a
      function of order: eigenvector centrality.} This panel is in the
    same format as Panel a, but here eigenvector centrality has been
    used to project the high-dimensional patterns of dynamic
    correlations onto a lower-dimensional space.
    \textbf{d.~Normalized timepoint decoding accuracy as a function of order:
      eigenvector centrality.} This panel is in the same format as
    Panel b, but here eigenvector centrality has been used to project
    the high-dimensional patterns of dynamic correlations onto a
    lower-dimensional space. See Figures~\kernelshape~and~\kernelwidth~for decoding results broken down by kernel shape and width, respectively.}
  \label{fig:decoding}
\end{figure}

@@

\begin{figure}[tp]
\centering
\includegraphics[width=1\textwidth]{figs/stats_heatmaps}
\caption{\textbf{Statistical summary of decoding accuracies for
    different neural features.}  Each column of matrices displays decoding
  results for one experimental condition (intact, paragraph, word, and
  rest).  We considered dynamic activity patterns (order 0) and
  dynamic correlations at different orders (order $> 0$).  We used
  two-tailed $t$-tests to compare the
  distributions of decoding accuracies obtained using each pair of
  features.  The distributions for each feature reflect the set of
  average decoding accuracies (across all kernel parameters), obtained
  for each random assignment of training and test groups.
  In the upper triangles of each matrix, warmer colors (positive $t$-values) indicate that the
  neural feature indicated in the given row yielded higher accuracy than the
  feature indicated in the given column.  Cooler colors (negative
  $t$-values) indicate that
  the feature in the given row yielded lower decoding accuracy than
  the feature in the given column.  The lower triangles of each map
  denote the corresponding $p$-values for the $t$-tests.  The diagonal
  entries display the relative average optimized weight given to each type of feature in
  a decoder that included all feature types (see \textit{Feature
    weighting and testing}).}
\label{fig:ttests}
\end{figure}

Both sets of temporal decoding analyses yielded qualitatively similar
results for the auditory (non-rest) conditions of the experiment
(Fig.~\ref{fig:decoding}: pink, green, and teal lines;
Fig.~\ref{fig:ttests}: three leftmost columns).  The highest
decoding accuracy for participants who listened to the intact
(unscrambled) story was achieved using high-order dynamic correlations
(PCA: second-order; eigenvector-centrality: fourth-order).  Scrambled
versions of the story were best decoded by lower-order correlations
(PCA/paragraph: first-order; PCA/word: order zero; eigenvector
centrality/paragraph: order zero; eigenvector centrality/word: order
zero).  The two sets of analyses yielded different decoding results on
resting state data (Fig.~\ref{fig:decoding}: purple lines;
Fig.~\ref{fig:ttests}: rightmost column).  We note
that, while the resting state times could be decoded reliably, the
accuracies were only very slightly above chance.  We speculate that
the decoders might have picked up on attentional drift, boredom, or
tiredness; we hypothesize that these all increased throughout the resting
state scan.  The decoders might be picking up on aspects of these
loosely defined cognitive states that are common across individuals.
The PCA-based approach achieved the highest resting state decoding
accuracy using order zero features (non-correlational,
activation-based), whereas the eigenvector centrality-based approach
achieved the highest resting state decoding accuracy using
second-order correlations.  Taken together, these analyses indicate
that high-level cognitive processing (while listening to the intact
story) is reflected in the dynamics of high-order correlations in
brain activity, whereas lower-level cognitive processing (while
listening to scrambled versions of the story that lack rich meaning)
is reflected in the dynamics of lower-order correlations and
non-correlational activity dynamics.  Further, these patterns are
associated both with the underlying activity patterns (characterized
using PCA) and also with the changing relative positions that
different brain areas occupy in their associated networks
(characterized using eigenvector centrality).

\begin{figure}[tp]
  \centering
  \includegraphics[width=\textwidth]{figs/most_abs}
  \caption{\textbf{Top terms associated with the most strongly
      correlated nodes at each order.}  Each color corresponds to one order of
    inter-subject functional correlations.  To calculate the dynamic
    correlations, eigenvector centrality has been used to project
    the high-dimensional patterns of dynamic correlations onto a
    lower-dimensional space at each previous order, which allows us
    the map the brain regions at each order by retaining the
    features of the original space. The inflated brain plots
    display the locations of the endpoints of the 10 strongest
    (absolute value) correlations at each order, thresholded at 0.999,
    and projected onto the cortical surface~\citep{CombEtal19}.  The
    lists of terms on the right display the top five Neurosynth
    terms~\citep{RubiEtal17} decoded from the corresponding brain maps
    for each order.  Each row displays data from a different
    experimental condition.  Additional maps and their corresponding
    Neurosynth terms may be found in the \textit{Supplementary
      materials} (intact: Fig.~\intact; paragraph: Fig.~\para; word:
    Fig.~\word; rest: Fig.~\rest).}
  \label{fig:neurosynth}
\end{figure}

Having established that patterns of high-order correlations are
informative to decoders, we next wondered which specific networks of
brain regions contributed most to these patterns.  As a representative
example, we selected the kernel parameters that yielded decoding
accuracies that were the most strongly correlated (across conditions
and orders) with the average accuracies across all of the
kernel parameters we examined.  Using Figure~\ref{fig:decoding}c as a
template, the best-matching kernel was a Laplace kernel with a width
of 50 (Fig.~\ref{fig:kernels}d; also see Fig.~\pca).  We used this kernel to compute a
single $K$ by $K$ $n^\mathrm{th}$-order DISFC matrix for each
experimental condition.  We then used Neurosynth~\citep{RubiEtal17} to
compute the terms most highly associated with the most strongly
correlated pairs of regions in each of these matrices
(Fig.~\ref{fig:neurosynth}; see \textit{Reverse inference}).

For all of the story listening conditions (intact, paragraph, and
word; top three rows of Fig.~\ref{fig:neurosynth}), we found that
first- and second-order correlations were most strongly associated
with auditory and speech processing areas.  During intact story
listening, third-order correlations reflected integration with visual
areas, and fourth-order correlations reflected integration with areas
associated with high-level cognition and cognitive control, such as
the ventrolateral prefrontal cortex.  However, when participants
listened to temporally scrambled stories, these higher-order
correlations instead involved interactions with additional regions
associated with speech and semantic processing (second and third rows
of Fig.~\ref{fig:neurosynth}).  By contrast, we found
a much different set of patterns in the resting state data
(Fig.~\ref{fig:neurosynth}, bottom row).
First-order resting state correlations were most strongly associated
with regions involved in counting and numerical understanding.
Second-order resting state correlations were strongest in visual
areas; third-order correlations were strongest in task-positive areas;
and fourth-order correlations were strongest in regions associated
with autobiographical and episodic memory.  We carried out analogous
analyses to create maps (and decode the top associated Neurosynth
terms) for up to fifteenth-order correlations (Figs.~\intact, \para,
\word, and \rest).  Of note, examining fifteenth-order correlations
between 700 nodes using conventional methods would have required
storing roughly
$\frac{700^{2 \times 15}}{2} \approx 1.13 \times 10^{85}$ floating
point numbers-- assuming single-precision (32 bits each), this would
require roughly 32 times as many bits as there are molecules in the
known universe!  Although these fifteenth-order correlations do appear
(visually) to have some well-formed structure, we provide this latter
example primarily as a demonstration of the efficiency and scalability
of our approach.

@@

The notion that cognition is reflected in (and possibly mediated by)
patterns of first-order network dynamics has been suggested by or
proposed in myriad empirical studies and
reviews~\citep[e.g.,][]{DemeEtal19, Turk13, LuriEtal18, FongEtal19,
  ParkEtal18b, PretEtal17, MannEtal18, RoyEtal19, LiegEtal19, ZouEtal19,
  ChanGlov10, GonzEtal19}.  Our study extends this line of work by
finding cognitively relevant \textit{higher-order} network dynamics
that reflect ongoing cognition.  Our findings also complement other work
that uses graph theory and topology to characterize how brain networks
reconfigure during cognition~\citep[e.g.,][]{BassEtal06, ZhenEtal19,
  McInJirs19, TokeSomm19, SizeEtal18, ReimEtal17, BetzEtal19}.

An open question not addressed by our study pertains to how different
structures integrate incoming information with different time
constants.  For example, one line of work suggests that the cortical
surface comprises a structured map such that nearby brain structures
process incoming information at similar timescales.  Low-level sensory
areas integrate information relatively quickly, whereas higher-level
regions integrate information relatively slowly~\citep{BaldEtal17,
  HassEtal08, HassEtal15, HoneEtal12a, LernEtal11, LernEtal14,
  ChieHone19}.  A similar hierarchy appears to play a role in
predicting future events~\citep{LeeEtal20}.  Other related work in
human and mouse brains indicates that the temporal response profile of
a given brain structure may relate to how strongly connected that
structure is with other brain areas~\citep{FallEtal19}.  Further study
is needed to understand the role of temporal integration at different
scales of network interaction, and across different anatomical
structures. Importantly, our analyses do not speak to the
physiological basis of higher-order dynamics, and could reflect
nonlinearities, chaotic patterns, non-stationarities, and/or
multistability, etc. However, our decoding analyses do indicate that
higher-order dynamics are consistent across individuals, and therefore
unlikely to reflect non-stimulus-driven dynamics that are unlikely to
be similar across individuals.

@@

Identifying high-order network dynamics associated with high-level
cognition required several important methods advances.  First, we used
kernel-based dynamic correlations to extended the notion of (static)
inter-subject functional connectivity~\citep{SimoEtal16} to a dynamic
measure of inter-subject functional connectivity (DISFC) that does not
rely on sliding windows~\citep[e.g., as in][]{MannEtal18}, and that
may be computed at individual timepoints.  This allowed us to
precisely characterize stimulus-evoked network dynamics that were
similar across individuals.  Second, we developed a computational
framework for efficiently and scalably estimating high-order dynamic
correlations.  Our approach uses dimensionality reduction algorithms
and graph measures to obtain low-dimensional embeddings of patterns of
network dynamics.  Third, we developed an analysis framework for
identifying robust decoding results by carrying out our analyses using
a range of parameter values and identifying which results were robust
to specific parameter choices. By showing that high-level cognition is
reflected in high-order network dynamics, we have elucidated the next
step on the path towards understanding the neural basis of cognition.
